---
title: "CodeBook"
author: "Mateusz Zajac"
date: "20.03.2015"
output: 
  html_document:
    keep_md: true
---

##Reading labels for files:
```{r}
labels <- read.csv("UCI HAR Dataset/features.txt", sep="", header = FALSE)
str(labels)
```

#Reading original data files
##Traning
###Read traning data and assign labels
```{r}
dataTrain <- read.csv("UCI HAR Dataset/train/X_train.txt", sep = "", header = FALSE, col.names = labels[,2])
```
###load and assign test subjects
```{r}
testSubjects <- read.csv(file = "UCI HAR Dataset/train/subject_train.txt", sep = "", header = FALSE, col.names = "subject")
dataTrain$subject <- testSubjects[,1]
```
###load and assign traning type labels
```{r}
traningType <- read.csv(file = "UCI HAR Dataset/train/y_train.txt", sep = "", header = FALSE, col.names = "label")
dataTrain$label <- traningType[,1]
```

##Test
###Read traning data and assign labels
```{r}
dataTest <- read.csv("UCI HAR Dataset/test/X_test.txt", sep = "", header = FALSE, col.names = labels[,2])
```
###load and assign test subjects
```{r}
testSubjects <- read.csv(file = "UCI HAR Dataset/test/subject_test.txt", sep = "", header = FALSE, col.names = "subject")
dataTest$subject <- testSubjects[,1]
```
###load and assign traning type labels
```{r}
traningType <- read.csv(file = "UCI HAR Dataset/test/y_test.txt", sep = "", header = FALSE, col.names = "label")
dataTest$label <- traningType[,1]
```

#Description of progress
At this point files are loaded into R. They are labeled properly. Next is merging and cleaning enviroment variables.

```{r echo=FALSE} 
rm(testSubjects)
rm(traningType)
```

##Merge - add rows
```{r}
names <- seq(from = nrow(dataTrain) + 1, length.out = nrow(dataTest))
rownames(dataTest) <- make.names(names, unique = TRUE)
names <- seq(from = nrow(dataTrain), length.out = nrow(dataTrain))
rownames(dataTrain) <- make.names(names, unique = TRUE)
data <- rbind(dataTrain, dataTest)
```

```{r echo=FALSE}
rm(names)
rm(dataTest)
rm(dataTrain)
rm(labels)
```

#Build sequence of columns to extract
```{r}
nseq <- c()
for(i in seq(as.integer(563 / 40))) { 
  nseq <- c(nseq, seq(from = ((i - 1) * 40 + 1), length.out = 6))
}

selectedData <- data[,c(nseq,562,563)]
selectedData$label <- as.character(selectedData$label)
```

#load activity labels
```{r}
activityLabels <- read.csv(file = "UCI HAR Dataset/activity_labels.txt", header = FALSE, sep = "")
activityLabels$V2 <- as.character(activityLabels$V2)
```

#replace label numbers with names
```{r}
apply(activityLabels, 1, function(row) {selectedData$label[selectedData$label == as.character(row[1])] <<- row[2]})
```

```{r echo=FALSE}
rm(nseq)
rm(i)
```
##Calculate the average of each variable for each activity and each subject.

```{r message=FALSE, warning=FALSE}
calculatedData <- selectedData[0,]
for(i in 1:30) {
  temp <- selectedData[selectedData$subject == i,]
  for(j in activityLabels$V2) {
    tempj <- temp[temp$label == j,]
    if(nrow(tempj) > 0) {
      tempj <- tempj[,1:84]
      tempj <- colMeans(x = tempj)
      tempj$label <- j; tempj$subject <- i;
      calculatedData <- rbind(calculatedData, tempj)
      calculatedData$label <- as.character(calculatedData$label)
    }
  }
}
```

##Save to file
```{r eval=FALSE, echo=FALSE}
write.table(x = calculatedData, file = "data.txt", sep = " ", row.names = FALSE)
```

#Calculated fields
At the end we get a mean for each observation for each subject and its activity type.

```{r echo=FALSE}
head(calculatedData, n=1)
tail(calculatedData, n=1)
```